---
title: "Lab 3 Homework"
author: Jonathan marquez 
date: "Winter 2019"
output:
  html_document:
    keep_md: yes
    theme: spacelab
---

```{r}
library(tidyverse)
```

```{r}
fisheries <- 
  readr::read_csv(file = "FAO_1950to2012_111914.csv") 
getwd()
```

```{r}
spec(fisheries)
```
1. Do you see any potential problems with the column names? Does the error message now make more sense?
Column names are duplicated so R manipulated them to unique names

```{r}
names(fisheries) ## the years are listed as separate columns themselves which is NOT what we want. 
```

2. The make.names() command is helpful when there are issues with column names. Notice that although the names are still cumbersome, much of the problemtatic syntax is removed.
```{r}
names(fisheries) = make.names(names(fisheries), unique=T) #changes the column names
names(fisheries) ## we'll see that the years now have a variable "x" in front of it. 
```

3. Let’s rename the columns. Use rename() to adjust the names as follows. Double check to make sure the rename worked correctly. Make sure to replace the old fisheries object with a new one so you can keep the column names.
country = Country..Country.
commname = Species..ASFIS.species.
sciname = Species..ASFIS.species._2
spcode = Species..ASFIS.species._1
spgroup = Species..ISSCAAP.group.
spgroupname = Species..ISSCAAP.group._1
region = Fishing.area..FAO.major.fishing.area.
unit = Measure..Measure.
```{r}
fisheries <-
fisheries %>%
  rename(
 country = Country..Country.,  ## I was having problems running the code because it turns out I already ran it once so I couldn't run it again. 
      commname    = Species..ASFIS.species., 
        sciname     = Species..ASFIS.species._2, 
     spcode      = Species..ASFIS.species._1, 
      spgroup     = Species..ISSCAAP.group., 
     spgroupname = Species..ISSCAAP.group._1, 
     region      = Fishing.area..FAO.major.fishing.area., 
     unit        = Measure..Measure.)
fisheries
```


4. Are these data tidy? Why or why not, and, how do you know? No, the column names are values of the variable date.
```{r}
glimpse(fisheries)## Looking at the data below we can determine it ins't tidy. Each year is listed as its own column and acts as a variable which is really problematic. 
```

5. We need to tidy the data using gather(). The code below will not run. I have added a bit of code that will prevent you from needing to type in each year from 1950-2012 but you need to complete the remainder.
```{r}
fisheries_tidy <- 
  fisheries %>% 
  gather(num_range('X',1950:2012), key='year', value='catch') ## we know have the year gathered in a single column 
fisheries_tidy ## the gather function created a "year" column as the key that's specifcally assigned to a  corresponding "value"

```

6. Use glimpse() to look at the categories of the variables. Pay particular attention to year and catch. What do you notice?
There are a lot of missing entries in the data
```{r}
glimpse(fisheries_tidy) ## Using glimpse, I can see that the year column is listed as a "chr" instead of "dbl". We also need to change blank values into NAs
```

7. From question 6 you should see that there are a lot of entries that are missing. In R, these are referred to as NA's but they can be coded in different ways by the scientists in a given study. In order to make the data tidy, we need to deal with them. As a preview to our next lab, run the following code. It removes the 'X' from the years and changes the catch column from a character into a numeric. This forces the blank entries to become NAs. The error "NAs introduced by coercion" indicates their replacement.
```{r}
fisheries_tidy <- 
  fisheries_tidy %>% 
  mutate(
    year= as.numeric(str_replace(year, 'X', '')), #remove the X from year. This will help us categorize it as a "chr"
    catch= as.numeric(str_replace(catch, c(' F','...','-'), replacement = '')) #replace character strings with NA
    )
```

8. Are the data tidy? Why?
The dates are organized into a new column called year and each observation has its own row.
```{r}
fisheries_tidy
```

9. You are a fisheries scientist studying cephalopod catch during 2012. Identify the top five consumers (by country) of cephalopods (don't worry about species for now). Restrict the data frame only to our variables of interest.
```{r}
fisheries_tidy %>%
  select(sciname, country, year, catch) %>%
  filter( year==2012, sciname=="Cephalopoda")%>% 
  arrange(desc(catch))

```

Top 5: India, Madagascar, Mozambique and Taiwan province of China 












