---
title: "HW 5"
author: "Jonathan Marquez "
date: "Winter 2019"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    theme: spacelab
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

Load the tidyverse
```{r}
library(tidyverse)
getwd()

```

library(tidyverse)
Mammals Life History
Let’s revisit the mammal life history data to practice our ggplot skills. Some of the tidy steps will be a repeat from the homework, but it is good practice. The data are from: S. K. Morgan Ernest. 2003. Life history characteristics of placental non-volant mammals. Ecology 84:3402.

1.Load the data.
```{r}
mammals <- readr::read_csv("mammal_lifehistories_v2.csv", na = c("", " ", "NA", "#N/A", "-999"))  ##make sure to put data in "" . won't work otherwise
```


2. Use your preferred function to have a look. Do you notice any problems?
```{r}
glimpse(mammals)
```


3. There are NA’s. How are you going to deal with them?
```{r}
mammals <-mammals %>% ## changed all "-999" into NAs
  na_if("-999")

```


4. Where are the NA’s? This is important to keep in mind as you build plots.
```{r}
mammals %>%
  purrr::map_df(~sum(is.na(.))) %>% 
  tidyr:: gather(variables,num_nas ) %>% # made a new data frame using gather from our already established variables and num_nas
  arrange(desc(num_nas))
```


5. Some of the variable names will be problematic. Let’s rename them here as a final tidy step.
```{r}
mammals <- 
  mammals %>%  
  dplyr::rename(
          genus        = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )
mammals
```



ggplot()
For the questions below, try to use the aesthetics you have learned to make visually appealing and informative plots. Make sure to include labels for the axes and titles.

options(scipen=999) #cancels the use of scientific notation for the session
6. What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship.
```{r}
options(scipen=999) #cancels the use of scientific notation for the session
mammals %>%
  ggplot(aes(x=newborn,y=gestation )) +
  geom_jitter()
```

7. You should notice that because of the outliers in newborn mass, we need to make some changes. We didn’t talk about this in lab, but you can use scale_x_log10() as a layer to correct for this issue. This will log transform the y-axis values.
```{r}
mammals %>%
  ggplot(aes(x=newborn,y=gestation )) +
  geom_jitter() +
  scale_x_log10() ## added as an additional layer 
```

8. Now that you have the basic plot, color the points by taxonomic order.
```{r}
mammals %>%
  ggplot(aes(x=newborn,y=gestation, color=order )) + ## added color with"color=x" colors by different groups!
  geom_jitter() +
  scale_x_log10() 
```

9. Lastly, make the size of the points proportional to body mass.
```{r}
mammals %>%
  ggplot(aes(x=newborn,y=gestation, color=order, size= mass )) + # added proportional sizes using "size=x" 
  geom_jitter() +
  scale_x_log10() 
```


10. Make a plot that shows the range of lifespan by order.
```{r}
mammals %>%
  ggplot(aes(x=order, y=max_life, fill=order)) + ## would it be wrong to switch X and Y?
  geom_boxplot() + 
  coord_flip()
```

