---
title: " HW 4"
author: "Jonathan Marquez"
date: "Winter 2019"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

Load the tidyverse
library(tidyverse)
```{r}
library(tidyverse)
getwd()
```

Practice and Homework
We will work together on the next part (time permitting) and this will end up being your homework. Make sure that you save your work and copy and paste your responses into the RMarkdown homnework file.

Aren’t mammals fun? Let’s load up some more mammal data. This will be life history data for mammals. The data are from: S. K. Morgan Ernest. 2003. Life history characteristics of placental non-volant mammals. Ecology 84:3402.

life_history <- readr::read_csv("data/mammal_lifehistories_v2.csv")
##Loading Data
```{r}
life_history <- readr::read_csv("mammal_lifehistories_v2.csv")
```


Rename some of the variables. Notice that I am replacing the old life_history data.
## Renaming some of the variables
```{r}
life_history <- 
  life_history %>% 
  dplyr::rename(
          genus        = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year` )
names(life_history)
glimpse(life_history)
```


Explore the data using the method that you prefer. Below, I am using a new package called skimr. If you want to use this, make sure that it is installed.
## Installing "skimr" package
```{r}
#install.packages("skimr")
library(skimr)
life_history %>%
  skimr::skim()
```

## Are there any NA’s in the data? Does this seem likely?
```{r}
msleep %>% 
  summarize(number_nas= sum(is.na(life_history))) ## doesn't know that na's are listed as "-999.0"
```
Using the code above there's no NAs in our data;however, the code was only programmed to specifically look for "NA"s when in reality NAs can be input under different characters/Values


Are there any missing data (NA’s) represented by different values? How much and where? In which variables do we have the most missing data? Can you think of a reason why so much data are missing in this variable?
## Yes, there is missing data. When I used the glimpse function I was actually able to see NAs input as "-999.0" . I'll run a code that will allow me to change the "-999.0" values into "Na" in which then we can recount the NA. 

```{r}
life_history_na <- 
  life_history %>% 
  na_if("-999") # changed "-999.0" to be recognized as "na"
life_history_na %>% 
  summarize(number_nas= sum(is.na(life_history)))

```
## Using the purr (a tidyverse function) we can map all the nas into a new data frame that will specifically show us 
```{r}
 Life_history_na <- ##I used both the "purr" and "gather" function to execute a beatufully data visual. changed to Cap L!!
 life_history_na %>%
 purrr::map_df(~ sum(is.na(.))) %>% 
  tidyr::gather(key="variables", value="num_nas") %>% 
  arrange(desc(num_nas))
Life_history_na
  


```
## Wean mass has the most missing information. It's tough distinguishing when animals wean completely sometimes. It's also tough being able to weigh some animals.Maybe many animals they observed didn't wean within their time of observation 


Compared to the msleep data, we have better representation among taxa. Produce a summary that shows the number of observations by taxonomic order.
## Lab 4_1 we were introduced to "group_by" function which is what i'll use to specify the number.  
```{r}
life_history %>%
group_by(order) %>%
  summarise(observations=n()) %>% ## we grouped by order and then further reduced our information to show how many observations we input for each order
  arrange(desc(observations))

```


Mammals have a range of life histories, including lifespan. Produce a summary of lifespan in years by order. Be sure to include the minimum, maximum, mean, standard deviation, and total n.
## I didn't understand this part fully 
```{r}
life_history %>% 
  mutate(lifespan=max_life/12) %>% ## added data to a new column to data. why did we divide by 12?
  group_by(order) %>%
  summarize(min=min(lifespan, na.rm=TRUE),
            max=max(lifespan, na.rm=TRUE),
            mean=mean(lifespan, na.rm=TRUE),
            sd=sd(lifespan, na.rm=TRUE),
            total=n()) 
 
  
```


Let’s look closely at gestation and newborns. Summarize the mean gestation, newborn mass, and weaning mass by order. Add a new column that shows mean gestation in years and sort in descending order. Which group has the longest mean gestation? What is the common name for these mammals?
```{r}
life_history%>%
  group_by(order) %>%
 summarize(mean_gestation=mean(gestation, na.rm=TRUE), ##when we summarize does that automatically reduce order?
            mean_newborn_mass=mean(newborn, na.rm=TRUE),
            mean_wean_mass=mean(wean_mass, na.rm=TRUE),
            total=n()) %>% 
  arrange(desc(mean_gestation))
  
  
  
```



















